---
title: '#Tidytuesday Cracking Passwords'
author: Brian Stephen Ssali
date: '2020-08-22'
slug: tidytuesday-cracking-passwords
categories: []
tags:
  - R
  - Tidytuesday
subtitle: ''
summary: ''
authors: []
lastmod: '2020-08-22T23:27:08+03:00'
featured: yes
image:
  caption: ''
  focal_point: ''
  preview_only: no
---

We are going to look a dataset from the Tidytuesday on the passwords. The dataset entails the actual text of the password,the category of the password,time taken to crack the password(both online and offline) and the strength of the password.

```{r setup, include=FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(collapse = TRUE,dpi = 200,cache = TRUE,cache.lazy = TRUE,tidy = "styler",fig.width = 8,fig.height = 8)
library(dplyr)
library(stringr)
library(ggplot2)
options("scipen"=999,digits = 4)

passwords <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-14/passwords.csv')

```


# Exploring the dataset
We use the skimr package to have a quick overview of the data with all the data types.

```{r overview_of_dataset ,message=FALSE,warning=FALSE}
library(skimr)
skimr::skim(passwords)
```

There are some rows with NAs that we are going to remove.It would be interesting to see the how long the passwords take to be cracked based on their corresponding categories and their strength as well

```{r glimpse_of_dataset ,message=FALSE,warning=FALSE,dpi=200}

plot1 <- passwords %>% filter(!is.na(category)) %>% mutate(time_unit = factor(time_unit,levels =  c("seconds","minutes","hours","days","weeks","months" ,"years"))) %>% 
  ggplot(aes(category,value,fill = time_unit))+
  geom_point(aes(colour = time_unit),size = 2.5)+
  coord_polar(clip = "on")+theme_dark()+
  theme(legend.position = "bottom",axis.text = element_text(colour = "black"),panel.grid.major = element_line("gray"))+ 
  scale_colour_manual(values = c("#FFDB6D","#C4961A","#F4EDCA","#D16103","#00AFBB","#4E84C4","#FC4E07"),guide = guide_legend(nrow = 1))

plot1

```

```{r strenght_of_pswd ,message=FALSE,warning=FALSE,dpi=200}
  passwords %>% filter(!is.na(category)) %>% 
    group_by(category) %>% transmute(strength = mean(strength,na.rm = T)) %>% 
    tidyr::pivot_longer(cols = c(strength)) %>% arrange(value) %>% distinct(.) %>% 
    ggplot(aes(value,reorder(category,value),fill = name))+
    geom_segment(aes(xend = 0,yend  = category))+
    geom_point(size = 4,aes(color = category),show.legend = F)+
    geom_text(aes(label = round(value,1)),hjust = -.5)+theme_minimal()+
    labs(x = "Strenght of Password",y = "Category of Password")
```

```{r nature_of_pwd,message=FALSE,warning=FALSE,dpi=200}

passwords %>% filter(!is.na(password)) %>% 
  mutate(num_digits = str_count(password,"\\d"),num_char = str_count(password,"\\D"),no_digits_char = str_count(password,"[^\\d] | [^\\D]"), 
         type_of_pwd = case_when(num_digits == 0 & num_char >0 ~"Character only",num_digits > 0 & num_char ==0 ~"Numeric only",num_digits >0 & num_char >0 ~"Both Character & Numeric")) %>%
  count(category,type_of_pwd,name = "Freq") %>% 
  ggplot(aes(Freq,category,fill = type_of_pwd))+
  geom_col(width = .5,position = "dodge")+theme_minimal()+
  theme(legend.position = "bottom")+labs(x = "Category of Password",y = "Frequecy")+
  scale_fill_manual(values = c("#00AFBB","#4E84C4","#FC4E07"))
```

Looks like majority of the passwords consist of only character values. Some of the password-related and sport related passwords have both the numeric and characters.The nerdy-pop passwords have the least number of passwords with numeric and simple-alphanumeric have the highest number of passwords with only numeric and also with both the numeric and characters. 

```{r message=FALSE,warning=FALSE,dpi=200}
  passwords %>% filter(!is.na(password)) %>% 
    mutate(num_digits = str_count(password,"\\d"),num_char = str_count(password,"\\D"),no_digits_char = str_count(password,"[^\\d] | [^\\D]"), 
           type_of_pwd = case_when(num_digits == 0 & num_char >0 ~"Character only",num_digits > 0 & num_char ==0 ~"Numeric only",num_digits >0 & num_char >0 ~"Both Character & Numeric"),
           labz = case_when(value > 75 ~ password,TRUE~"")) %>% 
    group_by(type_of_pwd,time_unit) %>% summarise(value = mean(value,na.rm = T)) %>% 
    mutate(time_unit = factor(time_unit,levels = c("seconds","minutes","hours","days","weeks","months" ,"years"))) %>% 
    ggplot(aes(type_of_pwd,value,fill = time_unit))+
    geom_col(position = "dodge")+theme_minimal()+
    theme(legend.position = "bottom")+
    labs(y = "Duration taken to crack password",x = NULL)+
    scale_fill_brewer(palette = "Set1",guide = guide_legend(nrow = 1))+scale_y_continuous(expand = c(0,0),limits = c(0,60))
```

Based on the data, its better to have a password that's alpha numeric as it has a higher password strength and would take longer to crack as compared to passwords that are only consisting of numeric or only characters.

